---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gdalptr

<!-- badges: start -->
<!-- badges: end -->

The goal of gdalptr is to learn about raw pointer handling and interface building from [adbcdrivermanager](https://github.com/paleolimbot/radbc/) and other fab projects by Dewey Dunnington. 


## Basic example

This is a basic example, first with R's C API (note use of   R_SetExternalPtrAddr in the source). 

```{r example}
library(gdalptr)
dsn <- system.file("extdata/volcano_gcp.tif", package = "gdalptr", mustWork  = TRUE)
xptr <- GDALOpen(dsn)
GDALGetRasterSize(xptr)  
```

(there was some work with the Rcpp wrapper left in branch 'rcpp-0.0.1'). 


We can have some fun with this, the dsn package has a few good online examples. 

```{r fun}
## remotes::install_github("hypertidy/dsn")

GDALGetRasterSize(GDALOpen(dsn::gebco22()))

GDALGetRasterSize(GDALOpen(dsn::wms_arcgis_mapserver_tms()))


```                  

The recursive VSI listing will tell us what files live in a tarball, a zip, a url endpoint, etc

```{r recursive}
tarball <-  pak::cache_list()$fullpath[1]
basename(tarball)
fs <- GDALVSIReadDirRecursive(file.path("/vsitar", tarball))
str(fs)
```



With an upcoming GDAL release (3.7), we can even target different aspects via GDAL's virtual driver, *without constructing actual VRT or opening the dataset*. 

```{r gdalvrt}
## because I'm working with the dev source
GDALVersion()

## we can target an overview in this massive tile dataset
GDALGetRasterSize(GDALOpen(sprintf("vrt://%s?ovr=12", dsn::wms_arcgis_mapserver_tms())))


GDALGetRasterSize(GDALOpen(sprintf("vrt://%s?ovr=6", dsn::wms_arcgis_mapserver_tms())))

GDALGetRasterSize(GDALOpen(sprintf("vrt://%s?ovr=0", dsn::wms_arcgis_mapserver_tms())))
```
